---
title: "Hands-on Exercise 2"
author: "Shaun Tan"
date: "4/25/2023"
date modified: "`r
Sys.Date()`"
format: html
execute:
    echo: true
    eval: true
    warning: false
    code-fold: true
    code-summary: "Show the code"
---

# Getting Started

## Installing and launching R packages

The code chunk below uses p_load() of pacman package to check if the required libraries are installed on the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse, ggrepel, ggthemes, hrbrthemes, patchwork)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```
## Examining the data

```{r}
summary(exam_data)
```

# Beyond ggplot2 Annotation: ggrepel

## Plotting the data using plain vanilla ggplot2

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## Plotting the data using ggrepel

### We simply replace geom_text() by geom_text_repel() and geom_label() by geom_label_repel.

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
# Beyond ggplot2 Themes

ggplot2 comes with eight built-in themes, they are: theme_gray(), theme_bw(), theme_classic(), theme_dark(), theme_light(), theme_linedraw(), theme_minimal(), and theme_void().

## Working with ggtheme package

### Plotting a histogram using theme_gray():

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### Plotting a histogram using theme_economist():

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### Plotting a histogram using theme_wsj():

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_wsj()
```

## Working with hrbtheme package

hrbrthemes package provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

### Plotting the distribution of Math scores using theme_ipsum()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

### Plotting the distribution of Math scores using theme_ipsum() but changing the text sizes and grid lines

1. axis_title_size argument is used to increase the font size of the axis title to 18,
2. base_size argument is used to increase the default axis label to 15, and
3. grid argument is used to remove the x-axis grid lines.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

# Beyond Single Graphs

It is not unusual that multiple graphs are required to tell a compelling visual story. There are several ggplot2 extensions provide functions to compose figure with multiple graphs. In this section, you will learn how to create composite plot by combining multiple graphs. First, let us create three statistical graphics by using the code chunk below.

First, we plot the histogram of the distribution of Math scores

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
```

Next, we plot the histogram of the distribution of English scores

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
```

Lastly, we will draw a scatterplot for English score versus Maths score by as shown below 

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## Creating Composite Graphics: pathwork methods

### Combining two ggplot2 graphs

```{r}
p1 + p2
```
### Combining three ggplot2 graphs

“|” operator to stack two ggplot2 graphs,
“/” operator to place the plots beside each other,
“()” operator the define the sequence of the plotting.

```{r}
(p1/p2) | p3
```
### Creating a composite figure with tag

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### Creating figure with insert

```{r}
p12 <- p1|p2
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

Creating a composite figure by using patchwork and theme_economist() of ggtheme

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```




