---
title: "Take-home Exercise 3 - MC3"
author: "Shaun Tan"
date: "6/3/2023"
date-modified: "`r Sys.Date()`"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
execute:
  warning: False
  echo: True
  eval: True
---


```{r}
pacman::p_load(jsonlite, tidygraph, ggraph, visNetwork, gralayouts, ggforce, tidytext, tidyverse)
```

## Importing json file by using jsonlite package

```{r}
mc3_data <-
  jsonlite::fromJSON("data/MC3.json")
```

```{r}
mc3_edges <- as_tibble(mc3_data$links)
```

```{r}
mc3_edges <-
  as_tibble(mc3_data$links) %>%
    distinct() %>%
  mutate(source = as.character(source),
         target = as.character(target),
         type = as.character(type)) %>%
  group_by(source, target, type) %>%
  summarise(weights = n()) %>%
  filter(source!=target) %>%
  ungroup()
``` 

```{r}
mc3_nodes <- as_tibble(mc3_data$nodes)
```

```{r}
mc3_nodes <- as_tibble(mc3_data$nodes) %>%
# distinct() %>%
  mutate(country = as.character(country),
         id = as.character(id),
         product_services = as.character(product_services),
         revenue_omu = as.numeric(as.character(revenue_omu)),
         type = as.character(type)) %>%
  select(id, country, type, revenue_omu, product_services)
```

## Initial 

need to recode the character(0) into NA to ensure future text sensing will be coherent

business group can be found by type (ie. company, beneficial owner, company contacts). beneficial owner will be connected to comapny. Company contacts will be connect to company. company connected to company 

look also at the description part. Some can be straightforward: automobile. Other can be computer product notebooks etc,

if you dont do text analytics, can just type in the filter and explore key words like "fish"