---
title: "In-class Exercise 4"
author: "Shaun Tan"
date: "5/6/2023"
date-modified: "`r Sys.Date()`"
execute:
  warning: False
  echo: True
  eval: True
#format:
  #html:
    #code-fold: true
    #code-summary: "Show the code"
---

```{r}
pacman::p_load(plotly, ggstatsplot, ggthemes, rstatix, gt, patchwork, tidyverse, png, webshot2)
```

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

::: panel-tabset

## The plot

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```

## The code chunk

```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
```
:::

# authoring with callout

::: {.callout-note}
We can see that the points deviate significantly from the straight diagonal line. this is a clear indication that the set of data is not normally distributed. 

:::

::: panel-tabset

## The plot

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
qq <- ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png    
```

## The code chunk

```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
qq <- ggplot(exam_data, 
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()

tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)

qq + table_png    
```

:::

remember to set.seed(1234) to ensure reproducibility
```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)
```

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "np",
  test.value = 60,
  xlab = "English scores"
)
```

since p value smaller than critical value (95% confidence level), reject the null hypothesis. 

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "p",
  messages = FALSE
)
```
