---
title: "In-class_Ex01"
---

# Getting Started

## Installing and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed on the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple horizontal bar chart

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar() +
  coord_flip() +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "lightblue", colour = "lightblue", 
                                    size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid', colour = "white"))
  labs(title = "Horizontal Histogram of Exam Data")
```

# Designing Data-drive Graphics for Analysis I

## The original design

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar() +
  labs(x = "Race", y = "Count", title = "Histogram of Exam Data")
```

Critique of this histogram:

1.  y-aixs label is not clear (i.e. count).
2.  The bars are notsorted by their resepctive frequencies.
3.  For static graph, frequency values should be added to provide addition information.

## The makeover design

```{r}
ggplot(data = exam_data,
       aes(x=reorder(RACE,RACE,
                     function(x)-length(x)), fill = RACE)) + 
  geom_bar() +
  ylim(0,220) +
  geom_text(stat="count", 
      aes(label=paste0(..count.., ", ", 
      round(..count../sum(..count..)*100, 1), "%")),
      vjust=-1) +
  labs(x = "Race", y = "No. of\nPupils") +
  theme_grey() +
  theme(axis.title.y=element_text(angle = 0))
```
# Designing Data-drive Graphics for Analysis II

## The original design
```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) + 
  geom_histogram(col = "#1b98e0",
                 bins=20) +
  ylim(0,45)
  
```
## The makeover design

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  geom_vline(aes(xintercept=mean(MATHS, na.rm=T)),
             color="red", 
             linetype="dashed", 
             size=1) +
  geom_vline(aes(xintercept=median(MATHS, na.rm=T)),
             color="grey30",
             linetype="dashed", 
             size=1)
  ylim(0,45)
```

# Designing Data-drive Graphics for Analysis III

## The original design

## The makeover design

```{r}
d <- exam_data   
d_bg <- d[, -3]  

ggplot(d, aes(x = ENGLISH, fill = GENDER)) +
  geom_histogram(data = d_bg, fill = "grey", alpha = .5) +
  geom_histogram(colour = "black") +
  facet_wrap(~ GENDER) +
  guides(fill = FALSE) +  
  theme_bw()
```
# Designing Data-drive Graphics for Analysis IV

## The original design

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point(size = 0.8) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

## The makeover design

```{r}
ggplot(data=exam_data, 
       aes(x=MATHS, y=ENGLISH)) +
  geom_point() +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  geom_hline(yintercept=50,
             linetype="dashed",
             color="grey60",
             size=1) + 
  geom_vline(xintercept=50, 
             linetype="dashed",
             color="grey60",
             size=1)

```